Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-03-29 bbbq_article.tex
Process feedback Geert, thanks @bogeert, progress https://github.com/richelbilderbeek/bbbq_article/issues/196

\richel{TODO: Tables don't have a legend, but instead have a title and footnotes below the table. Modify the tables to make it so in the end}
\richel{TODO: Condense figures into bigger ones with panels, following 1 message per figure. Order these figures chronologically}

@@

@@

For MHC-I, 9-mers were used, as this is the length most frequently presented in MHC-I and was used in our earlier
study (\cite{bianchi2017}). For MHC-II, 14-mers were used,  as these are the most frequently occurring
epitope length (\cite{bergseng2015different}).
We used a human (UniProt ID UP000005640\_9606), 
viral (SARS-CoV-2, UniProt ID UP000464024) 
and bacterial (\emph{Mycobacterium tuberculosis} (MTb), UniProt ID UP000001584) 
reference proteome. We used TMHMM  \richel{ref} to predict the topology 
of the proteins within these proteomes.
To predict the affinity of an epitope to a certain MHC molecule,
we used \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
and \verb;MHCnuggets; \richel{ref} for MHC-II.
The 13 MHC-I haplotypes used in this study are the same as 
used in our previous study (\cite{bianchi2017}).
The most frequent MHC-II haplotypes used  were selected, with a phenotypic frequency of at least 14\% in
the human population (\cite{greenbaum2011functional}),
resulting in 21 MHC-II haplotypes.
\richel{This is Frans' text, merge it with the current text: 
Unlike MHC-I where only one variable chain determines the landscape of the epitope binding pocket allowing for the classification of 5 HLA-A and 8 HLA-B superhaplotypes with .... variation in peptide binding repertoire, the MHC-II peptide binding groove is determined by both the alpha and beta chain. For HLA DQ both the alpha and beta chain is highly polymorphic  ..... \frans{why is HLA-DP not predicted?}
We took the alleles that represent percentage of the occurrence int he overall human population according to ... et al. 
}

This study differs in one important aspect from our previous work (\cite{bianchi2017}), 
The definition of a binder differs from \cite{bianchi2017}:
in the current study a peptide is called a binder if, within a certain haplotype, 
any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
the epitopes within a 
\emph{proteome} (see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2}
for values), whereas the original study defined
a binder as having an IC50 in the lowest 2\% 
of the peptides within a \emph{protein}.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
We believe that our revised definition precludes bias of proteins 
that give rise to no or only very few MHC epitopes

@@

The first query was a call to the 'gene' database for the 
term 'membrane protein' (in all fields) for the organism \emph{Homo sapiens}.
This resulted in 1,077 gene IDs.
% 1077 gene IDs is correct for December 2020.
% At 2021-03-01, one will get 1130 gene IDs.
% Also, one of the gene IDs that was valid back then,
% has been obsoleted.
The next query was a call to the 'gene' database 
to obtain the gene names from the gene IDs.
Per gene name, the 'dbSNP' NCBI database was queried for 
variations associated with the gene name.

@@

We recently predicted that peptides derived from TMHs of human proteins 
are over-presented on most haplotypes of HLA-A and HLA-B. Here, we wondered if this over-presentation is general 
and if peptides stemming from predicted TMHs of microbial pathogens 
would also be over-presented on HLA-A and HLA-B. We therefor analysed all superhaplotypes[GEERT: not explained, for instance in the Methods] of the HLA-A and HLA-B class (Figure \ref{fig:1}). Similar as for the human proteome, the percentages of MHC-I epitopes predicted to be overlapping with TMHs  was higher as expected by TMH abundance for the proteomes of SARS-CoV-2 and \emph{M. tuberculosis}.
See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) [GEERT what is in supplementary table 6? The numbers?]
The extent of the over-presentation between the various different HLA supertypes is similar for the probed proteomes, which strenghtens our previous conclusion [ref] that the hydrophopbicity of the MHC-binding groove is the main factor responsible for the predicted over-presentation of TMH-derived peptides.

\begin{figure}[!htbp]
  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
  \caption{
    Percentage of MHC-I epitopes predicted to be overlapping with TMHs
    for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
    The dashed horizontal lines indicate the percentage as expected by chance.
    See supplementary table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
  }
  \label{fig:1}
\end{figure}

@@

Additionally, we investigated if MHC-I and MHC-II are equally prone
to present TMH-derived epitopes. 
\richel{Add exact method}
See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}. [GEERT: Wat is er in 9 en 10? Wat is de conclusie? Dit moet nog wat verder uitgewerkt worden]

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation.png}
  \caption{
    Evolutionary conservation of SNPs in TMHs. [GEERT: legenad uitwerken. Wat is precies laten zien?]
  }
  \label{fig:conservation}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
  \caption{
    Evolutionary conservation of SNPs in TMHs,
    for either single- or multi-spanners/ [GEERT: ook uitwerken]
  }
  \label{fig:conservation_per_spanner}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
  \caption{
    Percentage of SNPs found in TMHs.
    Each point shows the predicted percentage of
    TMH (x-axis) and the observed occurrence of SNPs being located
    within a TMH (y-axis) for a protein.
    The dashed diagonal line shows the line of equality (i.e.
    equal conservation of TMHs and soluble protein regions).
    The red line [GEERT: where is the red line? Behind the blue line? That makes this useless and can be removed.]is a linear trend line that originates from the origin. 
    The blue line is a linear trend line that includes only the
    transmembrane proteins.
  }
  \label{fig:f_snps_found_and_expected}
\end{figure}

@@

\subsection{Elution studies}
[GEERT: ook hier: veel sterker is een korte conclusie, of boodschap]

@@

An important caveat of the elution studies is that we assume that the found epitopes are derived from their completely translated predecessor proteins.
However, two known biological phenomena that violate this assumption
are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
and peptide fusion (\cite{delong2016pathogenic}),
two processes that both result in peptide fragments
that did not originate from the mature proteins. [GEERT: mooi argument, maar moet wel kort uitgewerkt worden. Uitleggen wat DRiPs zijn en wat peptide fusion is, en hoe het kan leiden tot false positives]

Another caveat is that the TMH prediction tools are imperfect.
This imperfection, however, appears to be of limited importance,
as our data show that two different topology prediction software algorithms give similar results.

@@

\subsection{Presentation of epitopes derived from TMHs}

% \paragraph{MHC-II is expected to present TMHs}
\frans{we moeten iets schrijven over de verschillende haplotypen wat is the occurence van de haplotypen met andere worden hoe relevant zijn de haplotypen die veel TMH presenteren?}
The over-presentation of TMHs for MHC-II haplotypes 
is slight more pronounced than for MHC-I,
as with a third of the predicted binders 
\richel{Hier ben ik nu met verwerken...} [GEERT: dan stop ik hier eventjes? Of de rest ook weer keer lezen? Laat maar weten]
from the human and MtB proteome stem from TMH human reference proteome and MtB reference proteome to be presented on MHC-II. \frans{ik zoek nog naar associaties} 
such as \verb;HLA-DQA1*0101/DQB1*0501;

This study predict that TMH-derived epitopes are over-presented
in MHC-II as well. 
The intracellular pathway of MHC-II differs from that of MHC-I,
as it acquires epitopes from the exterior of the cell by phagocytosis.
How TMHs derived from exterior TMPs are lysed and loaded onto MHC-II 
is unknown. in vivo work is needed to elucidate this novel pathway.

% \paragraph{Immunogenicity}

In this study, we find that TMH-derived epitopes are over-presented.
The presentation of an epitope, however, 
is only a first step of an adaptive immune response.
We assume that the presentation of epitopes is
correlated with evoking a full immune response.
TMH-derived epitopes, however, are more hydrophobic than the average
epitope. If this evokes a stronger or weaker immune response is unknown,
and in vivo work is needed to resolve this.

\subsection{Evolutionary conservation of TMHs}

% \paragraph{Low number of SNPs}

The NCBI dbSNP database contains millions of SNPs.
As the NCBI API constrains its users to three calls per second
to assure fair use, we had to limit the extent of our analysis.
We assume we have taken a representative sample, but only
an analysis using the full NCBI database can confirm this.

@@

In general, on would hope that evolutionary selection results in
an immune system that as most attentive for loci that are
essential for a virus, as these will be most conserved.
In SARS-CoV-2, for example, there is preliminary evidence that the strongest
selection pressure is upon residues that changes its 
virulence (\cite{velazquez2020positive}).
These loci, however, only account for a small part of a pathogen's proteome.
Additionally, these essential parts differ widely between pathogens.
Because of this scarcity and variance in targets, 
one can imagine that the human immune system 
is not tailored to detect these sites, 
as hinted upon by an extensive study on influenza (\cite{han2019individual}),
where it was found that the selection pressure
exerted by the immune system was a non pro founding factor.

@@

TMHs, on the other hand, also have their functional constraints, 
and make up ... of a human, bacterial or viral proteome.
One can safely assume a proteome contains many TMHs,
therefore it may be beneficial for the host
if its immune system would be more attentive towards TMHs.
And maybe this has already happened: MHC-I already detects hydrophobic
peptides. This feature, however, may also be caused by selection
to detect hydrophobic regions in the soluble proteins of pathogens.
It is unknown, when focusing on TMHs only, if a signal of selection
can be detected.\frans{ik snap deze redenatie niet?}
We speculate that the human immune system is more attentive 
towards TMHs, as it is a general structure, that is evolutionarily conserved.

We aimed our evolutionary experiment at TMHs, because these can
be predicted well from a protein structure,
are common structures and are present in all pathogens. \frans{ dit doen we toch omdat we vinden dat er een preferred binding affinity is voor hydrophobe peptides?}
We could have done the same experiment on beta-sheets,
as also these can be predicted well (\cite{petersen2010netturnp}),
are common structures and are present in all pathogens.
We speculate that the human immune system may also be more attentive 
towards beta sheets for this reason.

@@

From this study, two important conclusions can be drawn. 
First, the MHC over-presentation of TMHs is likely a general feature 
and predicted to occur for most haplotypes of both MHC-I and -II 
and for humans as well as bacterial and viral pathogens. 
Second, TMHs are more evolutionary conserved that soluble protein motifs, 
at least in the human proteome.

% \paragraph{Over-presentation in vivo in MHC-I is robust}

We confirm that 1.4\% of epitopes that are naturally presented on MHC-I 
can be mapped to stem from a TMH. Additionally, this 
prediction is robust, as it holds for two TMH prediction tools, 
as shown by table \ref{tab:elution}).

% \paragraph{Over-presentation in vivo in MHC-II is robust}

We predict that also MHC-II presents epitopes derived from TMHs
in vivo, as we find that 3.9-4.0\% of sequenced epitopes
can be uniquely traced onto a human reference 
proteome (also shown in table \ref{tab:elution}).

Not only do we predict that MHC-I and MHC-II TMH-derived epitopes 
are presented in vivo, we conclude that these are additionally
presented more often than expected by chance 
alone (figures \ref{fig:1} and \ref{fig:2}).
This in silico finding is a general pattern 
that persists in both MHC-I and MHC-II, 
for humans, bacteria and viruses.

% \paragraph{Evolutionary conservation}

We conclude that mutations are more likely to occur
in soluble regions of proteins, than in TMHs.
When selecting SNPs at random, per 1000 SNPs found in a soluble domain, 
one finds on average 918 SNPs in TMHs.
This effect is more pronounced in multi spanners likely as these TMHs more often serve additional functions, besides spanning the lipid bilayer.

@@

We thank the Center for Information Technology of the University 
of Groningen for its support and for providing access to the Peregrine 
high performance computing cluster.

@@

RJCB and FB conceived the idea for this research. 
RJCB wrote the code.
RJCB and FB wrote the article.

@@

A part of this study does the same analysis as Bianchi et al., 2017.
Here we describe the deviations, which are about the use of different
software and the use of a different definition of what a binder is.

@@

The earlier study uses \verb;epitope-prediction; a hand-crafted method, 
that has been trained on MHC-I haplotypes only,
which is used here again. For MHC-II IC50 predictions, the
'mhcnuggetsr' R package is used.

@@

The earlier study defines a peptide a binder (for a haplotype), 
if \emph{within the peptide} in which it is found, 
is within the peptides with the 2\% lowest IC50 values.
This can be seen at \url{https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/master/predict-binders.R},
where the binders are written to file.

In this study, a peptide is defined a binder (for a haplotype), 
if within a \emph{proteome} in which it is found, 
is within the peptides with the 2\% lowest IC50 values,
where our previous study used the lowest 2\% IC50 values
within each \emph{protein}.
Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
for a binder per haplotype.

@@

Our previous study used the TMHMM web server
to predict TMHs.
The desktop version of TMHMM, however, gives an
error message on the 25 selenoproteins found in the human
reference proteome.
For the sake of reproducible research, we use the desktop version (as
we can call it from scripts) and, due to this, we removed the
selenoproteins from this analysis.

@@

For this research, the scientific literature was explored 
to identify the most recent free and open source (FOSS) prediction software.
This was done by searching for papers that (1) reference older
prediction software, and (2) present a novel method to make predictions.
As a starting point, a review paper was used.
For all software needed, we found no studies that compared contemporary tools 
in their prediction quality.

@@

There are multiple computational tools developed to predict which
parts of a protein forms a TMH.
In 2001, multiple of such prediction tools have been compared (\cite{moller2001evaluation}),
of which TMHMM (\cite{krogh2001predicting}) turned out to be the best, 
as is used in the previous study (\cite{bianchi2017}).
However, TMHMM has a restrictive software license and is nearly two
decades old.
Therefore, PureseqTM (\cite{wang2019efficient}),
was also used in this study, which has been more recently developed
and has a free software license.

@@

Also for MHC-II, there are multiple computational tools developed 
to predict epitopes,
such as using a trained neural network (\cite{nielsen2003reliable})
or a Gibbs sampling approach (\cite{nielsen2004improved}).
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
gave rise to the most accurate predictions.
The most recent and promising FOSS tool available now appears
to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
and MHC-II prediction, which we used for MHC-II predictions.

@@

To be able to call, from R, the TMH prediction 
software 'PureseqTM' (\cite{wang2019efficient}),
which is written in C, the package 'pureseqtmr' has been developed. 
'pureseqtmr' allows to install 'PureseqTM' and use most of its features.
Excluded are the features that are used by the 'PureseqTM' 
developers to verify the correctness of their work.
'pureseqtmr' has been submitted to and is accepted by CRAN.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Percentage of spots and spots that overlap with a TMH}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Table \ref{tab:f_tmh} shows an overview of the findings,
where a target specifies the source of the proteome,
where \verb;covid; denotes SARS-CoV-2 and \verb;covid; denotes
Mycobacterium tuberculosis. \verb;mhc_class; denotes the MHC
class, \verb;n_spots; the number of possible 9-mers (for MHC-I) 
or 14-mers (for MHC-II) possible. \verb;n_spots_tmh; the
number of epitopes that overlapped with a TMH that were binders. 
\verb;f_tmh; the percentage of peptides that had at least 1 residue
overlapping with a TMH.

@@

PureseqTM does not predict the topology
of proteins that have less than three amino acids. 
The TRDD1 ('T cell receptor delta diversity 1') protein,
however, is two peptides long. 
The R package \verb;pureseqtmr;, however, 
predicts that mono- and di-peptides are cytosolic.

>>>>

\richel{TODO: Tables don't have a legend, but instead have a title and footnotes below the table. Modify the tables to make it so in the end, \url{https://github.com/richelbilderbeek/bbbq_article/issues/181}}
\richel{TODO: Condense figures into bigger ones with panels, following 1 message per figure. Order these figures chronologically,
\url{https://github.com/richelbilderbeek/bbbq_article/issues/195}}

@@

\richel{I wonder if the order of the Methods, should match the improved
order of the Results as well, so first the MHC-I and MHC-II presentation,
followed by the experimental confirmation}

@@

For MHC-I, 9-mers were used, as this is the length most frequently presented in MHC-I and was used in our earlier
study (\cite{bianchi2017}). For MHC-II, 14-mers were used,  as these are the most frequently occurring
epitope length (\cite{bergseng2015different}).
We used a human (UniProt ID UP000005640\_9606), 
viral (SARS-CoV-2, UniProt ID UP000464024) 
and bacterial (\emph{Mycobacterium tuberculosis} (MTb), UniProt ID UP000001584) 
reference proteome. We used TMHMM  \richel{ref} to predict the topology 
of the proteins within these proteomes.
To predict the affinity of an epitope to a certain MHC molecule,
we used \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
and \verb;MHCnuggets; \richel{ref} for MHC-II.
The 13 MHC-I haplotypes used in this study are the same as 
used in our previous study (\cite{bianchi2017}).
The most frequent MHC-II haplotypes used  were selected, with a phenotypic frequency of at least 14\% in
the human population (\cite{greenbaum2011functional}),
resulting in 21 MHC-II haplotypes.

The MHC-I molecule presents epitopes of shorter length than MHC-II.
We expect this to be of no influence in the presentation level
of TMH-derived epitopes, for each of the three proteomes.
We verified this, by normalizing the percentage of TMH-derived epitopes found
to the percentage of TMH-derived epitopes expected, 
where a value of $1.0$ denotes that the percentage of TMH-derived epitopes found
matches the percentage as expected by chance.

Additionally, we investigated if MHC-I and MHC-II are equally prone
to present TMH-derived epitopes. 
\richel{Add exact method}
See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}.

In our previous work, we found that the observed pattern can be
explained from the hydrophobicity of the MHC-I binding 
cleft (\cite{bianchi2017}), we validated.
Here we do the same analysis for MHC-II \richel{Explain the analysis here}.

This study differs in one important aspect from our previous work (\cite{bianchi2017}), 
The definition of a binder differs from \cite{bianchi2017}:
in the current study a peptide is called a binder if, within a certain haplotype, 
any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
the epitopes within a 
\emph{proteome} (see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2}
for values), whereas the original study defined
a binder as having an IC50 in the lowest 2\% 
of the peptides within a \emph{protein}, 
which put less strain on the bioinformatics pipeline.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
\richel{Show side-by-side-comparison in supplementary materials}
We believe that our revised definition precludes bias of proteins 
that give rise to no or only very few MHC epitopes

@@

The first query was a call to the 'gene' database for the 
term 'membrane protein' (in all fields) for the organism \emph{Homo sapiens}.
This resulted in 1,077 gene IDs.
% 1077 gene IDs is correct for December 2020.
% At 2021-03-01, one will get 1130 gene IDs.
% Also, one of the gene IDs that was valid back then,
% has been obsoleted.
The next query was a call to the 'gene' database 
to obtain the gene names from the gene IDs.
Per gene name, the 'dbSNP' NCBI database was queried for 
variations associated with the gene name. 
As the NCBI API constrains its users to three calls per second
to assure fair use, we had to limit the extent of our analysis.

@@

Figure \ref{fig:1} shows the predicted presentation of TMH-derived peptides in MHC-I,
for a human, viral and bacterial proteome.
Per MHC-I haplotype, it shows the percentage of binders that overlap with a TMH 
for at least one residue.
A dashed line indicates the expected percentage of TMH-derived epitopes 
that would be presented, if TMH-derived epitopes would be presented just as 
likely as soluble regions.
In 12 out of 13 MHC-I haplotypes, TMH-derived epitopes are presented more often 
then the null expectation, for a human proteome. 
For the viral and bacterial proteome, 11 out of 13 haplotypes present
TMH-derived epitopes more often than expected by chance.
The extent of the over-presentation between the various different haplotypes 
is similar for the probed proteomes, which strenghtens our previous conclusion (\cite{bianchi2017}) that the hydrophopbicity of the MHC-binding groove is the main factor responsible for the predicted over-presentation of TMH-derived peptides.

\begin{figure}[!htbp]
  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
  \caption{
    Percentage of MHC-I epitopes predicted to be overlapping with TMHs
    for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
    The dashed horizontal lines indicate the percentage as expected by chance.
    See supplementary table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
    \richel{Add panel with hydrophobicity}
  }
  \label{fig:1}
\end{figure}

@@

Plotting the normalized values of presented TMH-derived epitopes (where
$1.0$ denotes that the percentage of presented TMH-derived epitopes
matches the values as expected by chance) in \ref{fig:rel_presentation}
shows that MHC-II has a higher relative over-presentation of TMH-derived
epitopes, when compared to MHC-I. \richel{HIERO}
The MHC-I molecule presents epitopes of shorter length than MHC-II.
We expect this to be of no influence in the presentation level
of TMH-derived epitopes, for each of the three proteomes.
We verified this, by normalizing the percentage of TMH-derived epitopes found
to the percentage of TMH-derived epitopes expected, 
where a value of $1.0$ denotes that the percentage of TMH-derived epitopes found
matches the percentage as expected by chance.
\richel{TODO: explain only the results found. Elaborate. Draw a conclusion}
Additionally, we investigated if MHC-I and MHC-II are equally prone
to present TMH-derived epitopes. 
\richel{Add exact method}
See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}.

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation.png}
  \caption{
    The number of SNPs in TMHs as expected by chance (left bar) 
    and found in the dbSNP database (right bar).
    Percentages are added to show the relative conservation
    of SNPs in TMHs found.
  }
  \label{fig:conservation}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
  \caption{
    The number of SNPs in TMHs as expected by chance (red bars) 
    and found in the dbSNP database (right bars), 
    after splitting the data into single- and multi-spanners.
    Percentages are added to show the relative conservation
    of SNPs in TMHs found.
  }
  \label{fig:conservation_per_spanner}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
  \caption{
    Percentage of SNPs found in TMHs.
    Each point shows the predicted percentage of
    TMH (x-axis) and the observed occurrence of SNPs being located
    within a TMH (y-axis) for a protein.
    The dashed diagonal line shows the line of equality (i.e.
    equal conservation of TMHs and soluble protein regions).
    The blue line is a linear trend line that includes only the
    transmembrane proteins.
  }
  \label{fig:f_snps_found_and_expected}
\end{figure}

@@

\subsection{TMH-derived epitopes are predicted to be presentated in vivo}

% \paragraph{Over-presentation in vivo in MHC-I and MHC-II is robust
We show that 1.4\% of epitopes that are naturally presented on MHC-I 
can be mapped to stem from a TMH. We also find that MHC-II presents epitopes derived from TMHs in vivo, as we find that 3.9-4.0\% of sequenced epitopes
can be uniquely traced onto a human reference 
proteome (also shown in table \ref{tab:elution}).

@@

An important caveat of the elution studies is that we assume that the found epitopes are derived from their completely translated predecessor proteins.
However,defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
and peptide fusion (\cite{delong2016pathogenic}) are
two processes that both result in peptide fragments
that did not originate from the mature proteins. [GEERT: mooi argument, maar moet wel kort uitgewerkt worden. Uitleggen wat DRiPs zijn en wat peptide fusion is, en hoe het kan leiden tot false positives]

Another caveat is that the TMH prediction tools are imperfect.
This imperfection, however, appears to be of limited importance,
as our data show that two different topology prediction software algorithms give similar results.
Additionally, this 
prediction is robust, as it holds for two TMH prediction tools, 
as shown by table \ref{tab:elution}).

@@

\subsection{TMH-derived epitopes are presented more often than expected by chance}

Not only do we predict that MHC-I and MHC-II TMH-derived epitopes 
are presented , we conclude that these are additionally
presented more often than expected by chance 
alone (figures \ref{fig:1} and \ref{fig:2}).
This in silico finding is a general pattern 
that persists in both MHC-I and MHC-II, 
for humans, bacteria and viruses.

% \paragraph{MHC-II is expected to present TMHs}
\frans{we moeten iets schrijven over de verschillende haplotypen wat is the occurence van de haplotypen met andere worden hoe relevant zijn de haplotypen die veel TMH presenteren?}
The over-presentation of TMHs for most MHC-II haplotypes 
is more pronounced than for MHC-I haplotypes,
as up to a third of the predicted MHC-II binders 
\richel{Hier ben ik nu met verwerken...} 
from the human and MtB proteome stem from TMHs. \frans{ik zoek nog naar associaties} 
such as \verb;HLA-DQA1*0101/DQB1*0501;

This results from this study support our hypothesis that TMH-derived epitopes are over-presented
in MHC-II, similarly to we previously showed for MHC-I [CITE BIANCHI].[GEERT: deel gewist, want was letterlijke herhaling van hierboven. Beter is zeggen dat (1) MHC-II presentatie belangrijk is voor B cell activatie. (2) B cellen antilichamen produceren, en dan antilichamen solvent-exposed protein regions herkennen. (3) Dat onze analyse suggereert HMC-II presentatie voornamelijk andere (TMH of buried) regio's betreft. (4) Dat T helper cellen dus andere delen van antigene eiwitten herkennen (MHC-II gepresenteerd door de B cellen) dan de BCR bindt.]

\subsection{Less SNPs in TMHs found, hinting at evolutionary conservation}

We find that mutations are more likely to occur
in soluble regions of proteins, than in TMHs.
When selecting SNPs at random, per 1000 SNPs found in a soluble domain, 
one finds on average 918 SNPs in TMHs.
This effect is more pronounced in multi spanners likely as these TMHs more often serve additional functions, besides spanning the lipid bilayer.

@@

In general, one would hope that evolutionary selection results in
an immune system that as most attentive for protein regions that are
essential for the survival, proliferation and/or virulence or pathogenic microbes, as these will be most conserved.
In SARS-CoV-2, for example, there is preliminary evidence that the strongest
selection pressure is upon residues that change its 
virulence (\cite{velazquez2020positive}).
These regions, however, only account for a small part of a pathogen's proteome.
Additionally, these essential parts might differ widely between different pathogenic proteins.
Because of this scarcity and variance in targets, 
one can imagine that it will be mostly unfeasible to provide innate immune responses against such essential protein regions, 
as hinted upon by a study on influenza (\cite{han2019individual}),
where it was found that the selection pressure
exerted by the immune system was a non profounding factor.

@@

TMHs, on the other hand, also have strict functional constraints as they need to span through hydrophobic lipid bilayers. Moreover, TMHs are abundant and make up ... of a human, bacterial or viral proteome.
Proteomes thus contain many TMHs,
and it might therefore be beneficial for host organisms
to tailor their immune systems towards immune responses against foreign TMHs.
This feature, however, may also be caused by selection
to detect hydrophobic regions in the soluble proteins of pathogens.
It is unknown, when focusing on TMHs only, if a signal of selection
can be detected.\frans{ik snap deze redenatie niet?}[GEERT: ik ook niet, vind beetje onnodig. Volstaat te zeggen dat TMHs niet de enige hydrofobe regio's zijn, en ook belang voor intra-protein en protein-protein interacties en catalytische sites. Daarnaast in deze paragraaf iets zeggen over single vs multi-TMH eiwitten]
We speculate that the human immune system is more attentive 
towards TMHs, as it is a general structure, that is evolutionarily conserved.

@@

From this study, two important conclusions can be drawn. 
First, the MHC over-presentation of TMHs is likely a general feature 
and predicted to occur for most haplotypes of both MHC-I and -II 
and for humans as well as bacterial and viral pathogens. 
Second, TMHs are more evolutionary conserved than solvent-exposed protein motifs, 
at least in the human proteome.

@@

We thank the Center for Information Technology of the University 
of Groningen for its support and for providing access to the Peregrine 
high performance computing cluster. GvdB is funded by a Young Investigator Grant from the Human Frontier Science Program (HFSP; RGY0080/2018), and a Vidi grant from the Netherlands Organization for Scientific Research (NWO-ALW VIDI 864.14.001). GvdB has received funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement No. 862137.

@@

RJCB and FB conceived the idea for this research. MVB helped with the proteome analysis of \emph{M. tuberculosis}.
RJCB wrote the code.
RJCB, FB and GvdB wrote the article.

@@

A part of this study does the same analysis as Bianchi et al., 2017.mainly concern the use of different
software and a different definition of what an MHC binder is.

@@

The earlier study used \verb;epitope-prediction; a hand-crafted method [GEERT: what is hand-crafted? They didn't use a computer? Why relevant?], 
that has been trained on MHC-I haplotypes only,
which is used here again. For MHC-II IC50 predictions, the
'mhcnuggetsr' R package was used in this study.

@@

The earlier study defined a peptide an MHC binder 
if \emph{within the protein} in which it was found, 
is was among the peptides with the 2\% lowest IC50 values.
This can be seen at \url{https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/master/predict-binders.R},
where the binders are written to file.

However, in this study, an MHC binder is defined as a peptide within a \emph{proteome} in which it is found, that is among the peptides with the 2\% lowest IC50 values.
Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
for a binder per haplotype. We believe that our revised definition is more correct, as it overcomes bias from proteins with very low numbers of peptides and/or MHC-predicted binders.

@@

Our previous study used the TMHMM web server
to predict TMHs.
The desktop version of TMHMM, however, gives an
error message on the 25 selenoproteins found in the human
reference proteome.
For the sake of reproducible research, we used the desktop version (as
we can call it from scripts) and, due to this, we removed the
selenoproteins from this analysis.

@@

For this research, we needed software to predict protein
topology, as well as the MHC-I and MHC-II binding affinities
of epitopes. We selected our software, by
searching the scientific literature 
to identify the most recent free and open source (FOSS) 
prediction software.
This was done by searching for papers that (1) cite older
prediction software, and (2) present a novel method to make predictions.
As a starting point, per type of prediction software,
a review paper was used (\cite{moller2001evaluation} for protein
topology, \cite{lundegaard2011prediction} for MHC-I
binding affinities and \cite{nielsen2003reliable} for MHC-II binding
affinities).

@@

There are multiple computational tools developed to predict which
parts of a protein forms a TMH.
In 2001, multiple of such prediction tools have been compared (\cite{moller2001evaluation}),
of which TMHMM (\cite{krogh2001predicting}) turned out to be the most accurate, 
as is used in the previous study (\cite{bianchi2017}).
However, TMHMM has a restrictive software license and is nearly two
decades old.
Therefore, PureseqTM (\cite{wang2019efficient}),
was also used in this study, which has been more recently developed
and has a free software license.

@@

Also for MHC-II, there are multiple computational tools developed 
to predict epitopes,
such as using a trained neural network (\cite{nielsen2003reliable})
or a Gibbs sampling approach (\cite{nielsen2004improved}).
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
made the most accurate predictions.
The most recent FOSS tool available now appears
to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
and MHC-II predictions. 
As we already use \verb;epitope-prediction; (\cite{bianchi2017}) 
for MHC-I predictions, we use MHCnuggets only for MHC-II predictions.

@@

To be able to call, from R, the TMH prediction 
software 'PureseqTM' (\cite{wang2019efficient}),
which is written in C, the package 'pureseqtmr' has been developed. 
'pureseqtmr' allows to install 'PureseqTM' and use most of its features.
'pureseqtmr' has been submitted to and is accepted by CRAN.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Prediction of percentage of epitopes overlapping with a TMH}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Table \ref{tab:f_tmh} shows an overview of the findings,
where a target specifies the source of the proteome,
where \verb;covid; denotes SARS-CoV-2 and \verb;myco; denotes
Mycobacterium tuberculosis. \verb;mhc_class; denotes the MHC
class, \verb;n_spots; the number of possible 9-mers (for MHC-I) 
or 14-mers (for MHC-II) possible. \verb;n_spots_tmh; the
number of epitopes that overlapped with a TMH that were binders. 
\verb;f_tmh; the percentage of peptides that had at least 1 residue
overlapping with a TMH.

@@

PureseqTM does not predict the topology
of proteins that have less than three amino acids. 
The TRDD1 ('T cell receptor delta diversity 1') protein,
however, is two amino acids long. 
The R package \verb;pureseqtmr;, however, 
predicts that mono- and di-peptides are cytosolic.
